Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-04-19 apj.tex
error bars on intrinsic covariance matrix

Non-trivial residual magnitude dispersions are captured in $C_c$.  The diagonal elements are captured by the $\sigma$ parameters;
the residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.  Given that the 
off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix rather than the matrix elements themselves,
it is not straightforward to present correlation matrix confidences:
the average over Monte Carlo links of the
Cholesky factors will not  generally yield a correlation matrix.  
To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
chain, element by element.
For $UBVRI$ the matrix is
\begin{equation}
C_c(U,B,V,R,I)=
\begin{pmatrix}
\begin{array}{rrrrr}
0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
\end{array}
 \end{pmatrix} \text{mag}^2.
 \label{mag_cov:eqn}
 \end{equation}
The average correlation matrix is
\begin{equation}
Cor(U,B,V,R,I)=
\begin{pmatrix}
\begin{array}{rrrrr}
1.00 & 0.42 & -0.18 & -0.05 & 0.12 \\
0.42 & 1.00 & 0.26 & -0.09 & -0.39 \\
-0.18 & 0.26 & 1.00 & 0.09 & -0.22 \\
-0.05 & -0.09 & 0.09 & 1.00 & 0.27 \\
0.12 & -0.39 & -0.22 & 0.27 & 1.00
\end{array}
 \end{pmatrix}.
 \end{equation}
The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
expressed as the standard deviations and
correlation matrix
 \begin{equation}
 \sigma(U-V, B-V, V-R, V-I)=
 \begin{pmatrix}
0.068 & 0.034 & 0.023 & 0.053
  \end{pmatrix} \text{mag},
 \label{color_sd:eqn}
 \end{equation}
 \begin{equation}
 Cor(U-V, B-V, V-R, V-I)=
\begin{pmatrix}
\begin{array}{rrrr}
1.000 & 0.602 & -0.353 & -0.317 \\
0.602 & 1.000 & -0.197 & 0.082 \\
-0.353 & -0.197 & 1.000 & 0.647 \\
-0.317 & 0.082 & 0.647 & 1.000
\end{array}
\end{pmatrix}.
  \label{color_cor:eqn}
 \end{equation}
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where direct comparison can be made.

@@

\section{Supernova Model II: Two Extrinsic Parameters, One Intrinsic Parameter}
\label{model2:sec}
In \S\ref{model:sec} we show that the SNfactory data analyzed with an agnostic model with
two extrinsic parameters produce
magnitudes consistent
with two-parameter dust models that have been established for stars in the Milky Way, LMC, and SMC.
An extension to our model is now added to explore the possibility of
a possible additional
color parameter that could have its origins in supernova astrophysics.
The extended model referred to as Model~II is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_U+\alpha_U EW_{Ca} + \beta_U EW_{Si} + \eta_U \lambda_{Si} + \delta_U D\\
c_B+\alpha_B EW_{Ca} + \beta_B EW_{Si} + \eta_B \lambda_{Si} + \delta_B D \\
c_V+\alpha_V EW_{Ca} + \beta_V EW_{Si} + \eta_V \lambda_{Si} + \delta_V D\\
c_R+\alpha_R EW_{Ca} + \beta_R EW_{Si} + \eta_R \lambda_{Si} + \delta_R D\\
c_I+\alpha_I EW_{Ca} + \beta_I EW_{Si}+ \eta_I \lambda_{Si} + \delta_I D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{U} k_0 +\gamma^1_{B} k_1 \\B +\gamma^0_{B} k_0 +\gamma^1_{B} k_1 \\
V+\gamma^0_{V} k_0+\gamma^1_{V} k_1\\R+\gamma^0_{R} k_0 + \gamma^1_{R} k_1\\I+\gamma^0_{I} k_0+\gamma^1_{I} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
In all aspects it is identical to the previous model except for the addition of the term $\delta D$,
which affects the mean
absolute magnitudes.  Thus
each supernova is given an intrinsic $D$ parameter that is linearly related to
absolute magnitudes through global coefficients $\delta$.  As with $k_0$ and $k_1$,  $D$ is a latent supernova parameter whose existence is inferred from
its effects on broadband magnitudes.

@@

The $\delta D$ terms contribute to the mean of the Normal distribution in Eq.~\ref{ewsiv:eqn}, together with the spectral-feature terms:
it is for this reason that for convenience these terms are referred to as ``intrinsic''.
However, we emphasize now that we have also run the alternative case  of switching the placement of the
$\delta D$ terms from Eq.~\ref{ewsiv:eqn} to Eq.~\ref{dust:eqn}, i.e.\ from intrinsic to extrinsic contributions.
Analyses of the two different models give practically identical results, meaning that data are not constraining enough
to identify $\delta D$ as being physically intrinsic or extrinsic.

>>>>

Non-trivial residual magnitude dispersions are captured in $C_c$.  The diagonal elements are captured by the $\sigma$ parameters;
the residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
\color{orange}
Of particular interest in $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in $V$ and colors with respect to $V$ are
\begin{equation}
\sigma(V,U-V,B-V,V-R,V-I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
\left(
0.021^{+0.004}_{-0.005} ,
0.067^{+0.010}_{-0.010},
0.034^{+0.005}_{-0.005},
0.023^{+0.002}_{-0.002},
0.053^{+0.004}_{-0.004}
\right)
%\end{array}
% \end{pmatrix} 
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{equation}
The correlations between these parameters are
\begin{equation}
Cor(V,U-V,B-V,V-R,V-I)=
\begin{pmatrix}
\begin{array}{rrrrr}
1 & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
-0.48^{+0.20}_{-0.18} & 1 & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
-0.34^{+0.21}_{-0.22} & 0.60^{+0.10}_{-0.13} & 1 & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
0.88^{+0.10}_{-0.25} & -0.38^{+0.18}_{-0.17} & -0.21^{+0.17}_{-0.17} & 1 & 0.65^{+0.06}_{-0.07} \\
0.60^{+0.11}_{-0.19} & -0.33^{+0.15}_{-0.14} & 0.08^{+0.14}_{-0.15} & 0.65^{+0.06}_{-0.07} & 1 \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{equation}

\color{black}
%
%Given that the 
%off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix rather than the matrix elements themselves,
%it is not straightforward to present correlation matrix confidences:
%the average over Monte Carlo links of the
%Cholesky factors will not  generally yield a correlation matrix.  
%To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
%chain, element by element.
%For $UBVRI$ the matrix is
%\begin{equation}
%C_c(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
%0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
%-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
%-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
%0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
%\end{array}
% \end{pmatrix} \text{mag}^2.
% \label{mag_cov:eqn}
% \end{equation}
%The average correlation matrix is
%\begin{equation}
%Cor(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%1.00 & 0.42 & -0.18 & -0.05 & 0.12 \\
%0.42 & 1.00 & 0.26 & -0.09 & -0.39 \\
%-0.18 & 0.26 & 1.00 & 0.09 & -0.22 \\
%-0.05 & -0.09 & 0.09 & 1.00 & 0.27 \\
%0.12 & -0.39 & -0.22 & 0.27 & 1.00
%\end{array}
% \end{pmatrix}.
% \end{equation}
%The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
%expressed as the standard deviations and
%correlation matrix
% \begin{equation}
% \sigma(U-V, B-V, V-R, V-I)=
% \begin{pmatrix}
%0.068 & 0.034 & 0.023 & 0.053
%  \end{pmatrix} \text{mag},
% \label{color_sd:eqn}
% \end{equation}
% \begin{equation}
% Cor(U-V, B-V, V-R, V-I)=
%\begin{pmatrix}
%\begin{array}{rrrr}
%1.000 & 0.602 & -0.353 & -0.317 \\
%0.602 & 1.000 & -0.197 & 0.082 \\
%-0.353 & -0.197 & 1.000 & 0.647 \\
%-0.317 & 0.082 & 0.647 & 1.000
%\end{array}
%\end{pmatrix}.
%  \label{color_cor:eqn}
% \end{equation}
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where direct comparison can be made.
\color{orange}
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eqn.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in the inferred slopes and smaller statistical uncertainties.
\color{black}

@@

\section{Supernova Model II: Two Extrinsic Parameters, One Intrinsic Parameter}
\label{model2:sec}
In \S\ref{model:sec} we show that the SNfactory data analyzed with an agnostic model with
two extrinsic parameters produce
magnitudes consistent
with two-parameter dust models that have been established for stars in the Milky Way, LMC, and SMC.
An extension to our model is now added to explore the possibility of
a possible additional
color parameter that could have its origins in supernova astrophysics.
The extended model referred to as Model~II is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_U+\alpha_U EW_{Ca} + \beta_U EW_{Si} + \eta_U \lambda_{Si} + \delta_U D\\
c_B+\alpha_B EW_{Ca} + \beta_B EW_{Si} + \eta_B \lambda_{Si} + \delta_B D \\
c_V+\alpha_V EW_{Ca} + \beta_V EW_{Si} + \eta_V \lambda_{Si} + \delta_V D\\
c_R+\alpha_R EW_{Ca} + \beta_R EW_{Si} + \eta_R \lambda_{Si} + \delta_R D\\
c_I+\alpha_I EW_{Ca} + \beta_I EW_{Si}+ \eta_I \lambda_{Si} + \delta_I D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv2:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{U} k_0 +\gamma^1_{B} k_1 \\B +\gamma^0_{B} k_0 +\gamma^1_{B} k_1 \\
V+\gamma^0_{V} k_0+\gamma^1_{V} k_1\\R+\gamma^0_{R} k_0 + \gamma^1_{R} k_1\\I+\gamma^0_{I} k_0+\gamma^1_{I} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust2:eqn}
\end{equation}
In all aspects it is identical to the previous model except for the addition of the term $\delta D$,
which affects the mean
absolute magnitudes.  Thus
each supernova is given an intrinsic $D$ parameter that is linearly related to
absolute magnitudes through global coefficients $\delta$.  As with $k_0$ and $k_1$,  $D$ is a latent supernova parameter whose existence is inferred from
its effects on broadband magnitudes.

@@

The $\delta D$ terms contribute to the mean of the Normal distribution in Eq.~\ref{ewsiv2:eqn}, together with the spectral-feature terms:
it is for this reason that for convenience these terms are referred to as ``intrinsic''.
However, we emphasize now that we have also run the alternative case  of switching the placement of the
$\delta D$ terms from Eq.~\ref{ewsiv2:eqn} to Eq.~\ref{dust2:eqn}, i.e.\ from intrinsic to extrinsic contributions.
Analyses of the two different models give practically identical results, meaning that data are not constraining enough
to identify $\delta D$ as being physically intrinsic or extrinsic.
